Methods

### 2021-09-09 text/main.Rmd
methods edits

---
output:
  word_document:
    fig_caption: yes
    reference_docx: ../text/word_template_science.docx
bibliography: ../text/global_proc.bib
csl: science.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
loadd(fig1)
loadd(fig2)
loadd(fig3)
loadd(fig4)
```

@@

**Abstract:**
Preserving the functioning of coral reefs is a critical challenge of the 21st century. However, a lack of quantitative assessments of multiple functions across large spatial scales has hindered local and regional conservation efforts. We integrate empirically-parametrized bioenergetic models and global community surveys to quantify five key functions mediated by coral reef fishes. We show that functions exhibit critical trade-offs driven by varying community structures, such that no reef can holistically maximize functioning. Further, functions are locally dominated by few species, but worldwide, the identity of dominant species greatly varies; 70% of the 1,110 species in our dataset are functionally dominant. Our results underline the need for a nuanced approach to coral reef conservation that considers multiple functions beyond the effect of standing stock biomass.

**Main Text:**
The flow of elements through biological communities fuels all ecosystems on Earth [@Welti2017].
Contemporary ecosystems and their functioning are threatened by the harmful dominance of humans [@Ceballos2015].
Coral reefs represent a prime example of an ecosystem that is severely impacted by anthropogenic threats including warming, and intensive fishing pressure [@Hughes2017; @Pauly2002].
Drastic declines in habitat quality and fish biomass coral reefs have evoked serious concerns about the persistence of coral reef functioning [@Bellwood2019; @Williams2019; @Brandl2019front).
However, our knowledge on process-based functions - fluxes of elements and energy - is still in its infancy and evaluations of coral reef functioning typically relies on static proxies, such as live coral cover, standing stock biomass of reef fishes, or the diversity of qualitative traits [e.g. @Cinner2020].
The sparsity of quantification of functions constitutes a severe limitation to efficient management [@Brandl2019front]: How can we conserve functions if we can not quantify them?

Coral reef fishes play an essential role in coral reef functioning, because coral reef fishes contribute to a high proportion of consumer biomass and therefore represent essential vectors of fluxes of carbon, nitrogen and phosphorus through storage and recycling (McIntyre et al. 2008; All-
geier et al. 2014; Barneche et al. 2014). Broad ecosystem functions mediated by coral reef fishes include nutrient cycling, biomass production, herbivory , and piscivory (secondary consumption) (Brandl et al. 2019front).
While the high diversity of coral reef fishes has inspired many studies with a focus on ecosystem functioning, only a handful of studies have attempted to quantify functions as continuous fluxes (Brandl et al. 2019a). Rather, most studies define fish-mediated ecosystem functioning with proxies such as trait diversity or total biomass (e.g. Mora& Sale 2011; e.g., Mouillot et al. 2014). Further, studies that do quantify functions are mostly focused on biomass production (e.g., @Morais2019; @Morais2020glob) or fish excretion (e.g., @Allgeier2014). An assessment of multiple functions on a large scale is currently lacking.
Therefore, trade-offs between functions, their drivers, and their vulnerability to global stressors remain poorly understood in coral reef ecosystems [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the elemental fluxes that underpin reef fish functioning.
Using empirically-collected species-specific data on basic organismal processes and Bayesian phylogenetic models, we parameterize individual-level bioenergetic models to estimate five key ecosystem functions: nitrogen (N) excretion, phosphorus (P) excretion, biomass production, herbivory, and piscivory.
We apply these bioenergetic models to 9,118 reef fish communities across 585 sites worldwide (Table S1) to:
(1) quantify community-level reef fish functions and their trade-offs,
(2) extract the community- and species-level effects on these functions, and
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

@@

Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we demonstrate a predictably strong relationship between fish biomass and all five functions by running multivariate mixed effect Bayesian model (Fig. 1). Specifically, the effects of the log-transformed biomass on log-transformed functions were 0.932 (95\%CI: 0.929; 0.934), 1.051 (1.047; 1.056), 0.771 (0.764; 0.780), 0.940 (0.923; 0.957), and 0.668 (0.635; 0.702) for N excretion, P excretion, production, herbivory, and piscivory, respectively. We also incorporated sea surface temperature (sst) because it affects the metabolic and growth rates of fishes, which scales up to the community [@Barneche2019]. We found positive effects of temperature on N excretion, production, and herbivory, and negligible effects of sst on P excretion and piscivory (suppl table).

The above-mentioned multivariate model also allowed us to estimate the correlations between functions, independent of the effects of biomass and sst.
Specifically, we estimated correlations between functions on three levels: the locality effects, the site effects, and the residual variations (Fig. 2, suppl. fig x).
The correlations mostly overlapped across the three levels for a given pair of functions. We found consistently negative trade-offs between P excretion and N excretion, P excretion and biomass production. Further, there were slightly weaker negative correlations between piscivory and N excretion and between piscivory and herbivory. As a consequence of the negative correlations, a fish community cannot have high values for all functions simultaneously.

@@

We then questioned if functions tend to be driven by specific species across sites. We quantified the degree of dominance inside each community on the site-level and found that functions are on average charachterized by a high level of dominance, i.e. in most communities, a small proportion of species perform most of a function (Fig. 4a). As a comparison, we simulated random communities ranging from 5 to 50 species which yielded an average XX. The predicted average of the degree of dominance was x, x, x, x, and x for each function respectively.
We also calculated the proportion of species that is dominant in at least one transect (i.e. species that contribute more than a community in which each species contributes equally). We found that most species play a dominant role, and 70\% of all species contributed disproportionally to a function in at least one community (Fig. 4b).
We finally investigate how often species are dominant across their range and found that very few species are dominant throughout their range (Fig. 4c). Together, this means that functions inside communities tend to be driven by few dominant species, but that the identity of these dominant species varies highly across communities.

@@

By quantifying five key functions mediated by coral reef fishes, we demonstrate that ecosystem-wide functioning is mainly driven by fish biomass, yet biomass alone does not fully capture functional variability across communities.
While biomass is one of the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], our analyses suggests that biomass alone does not sufficiently characterize functioning and using biomass as a proxy for functioning masks fundamental differences in critical community-level functions.
We demonstrate strong trade-offs among the five functions, independent of biomass.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize all functions.

Fish community structure drives the observed trade-offs among functions which can be explained by organismal physiology and life-history traits of the individuals inside communities [@Schiettekatte2020; @Schramski2015].
For example, we observed a clear trade-off between P excretion and biomass production which is largely driven by the age and trophic community structure.
Communities dominated by fishes with high trophic levels are charachterized high P excretion rates because predatory fishes have a P-rich diet (@Allgeier2014).
In contrast, biomass production is high in communities dominated by fishes with low trophic levels because herbivores tend to have higher growth rates [@Morais2018].
Moreover, communities dominated by mature fishes have high values of P excretion because growing fishes are more likely to retain P to form P-rich bones [@Schiettekatte2020; ref], while communities dominated by immature fishes logically maximize biomass production.
Furthermore, metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportionate contributions to functions that rely on rapid energetic turnover such as herbivory, piscivory, production, and N excretion [@Barneche2018; @Morais2020glob; @Brandl2019science].

Further, our results reveal that functions consistently hinge on a few dominant species, but the identity of local, dominant species vary strongly across sites [@Lefcheck2019].
Locally, specific high-performing taxa may disproportionately impact rates of functioning at the community level due to high biomass or abundance [@Topor2019].
Specifically, on average, more than 50% of a given function is upheld by only 12% of the species present within a local community.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
While few high-performing taxa indeed dominate functioning in each site, there are no species that are dominant across their entire range.
In addition, more than half of all species contributed disproportionally to a specific function in at least one site inside their range.

@@

Conserving biomass, diversity, and ecosystem functions are important objectives of most conservation initiatives [@Cinner2020].
While safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory.
However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards herbivore domination may result in the decline of P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions, yet our comprehension of
process-based functioning of coral reefs or what a “functional” coral reef really means is still in its infancy [@Brandl2019front].
To appropriately manage coral reef functioning with local needs and stressors, future research should seek to establish functional baselines for global coral reefs.
Until then, our study suggests that coral reef fish functions can be managed by maximizing standing stock biomass, protecting local key species, and promoting a balanced community structure.

@@

# Figures

@@

**References and Notes:**
<div id="refs"></div>

**Acknowledgements:**
We thank the staff at CRIOBE, Moorea for field support. We would also like to thank Jérémy Carlot, Beverly French, Titouan Roncin, Yann Lacube, Camille Gache, Gabrielle Martineau, Kailey Bissell, Benoit Espiau, Calvin Quigley, Kaitlyn Landfield and Tommy Norin for their help in the field, Sophie Schiettekatte for proof-reading the manuscript, and Guillemette de Sinéty and Jérémy Wicquart for their contribution to otolith analyis.
**Funding:**
This research was funded by the BNP Paribas Foundation (Reef Services Project) and the French National Agency for Scientific Research (ANR; REEFLUX Project; ANR‐17‐CE32‐0006). This research is product of the SCORE-REEF group funded by the Centre de Synthèse et d’Analyse sur la Biodiversité (CESAB) of the Foundation pour la Recherche sur la Biodiversité (FRB) and the Agence Nationale de la Biodiversité (AFB). VP was supported by the Institut Universitaire de France (IUF) and JMC was supported by a Make Our Planet Great Again Postdoctoral Grant (mopga‐pdf‐0000000144).
**Author contributions:**
NMDS and VP conceived the idea and NMDS, VP, SJB, and JMC designed methodology; NMDS, JMC, SJB, AM, FM, VP, KSM, JEA and DEB collected the data; All authors shared existing data. NMDS analyzed the data and led the writing of the manuscript. All authors contributed significantly to the drafts and approved the final version for publication.
**Competing interests:** None declared.
**Data and materials availability:**
All data and code to reproduce figures are available online through GitHub (https://github.com/nschiett/global_proc) and figshare (https://figshare.com/s/f789aec2c20492c4f0f9). All data on individual fish traits are available from the corresponding author on reasonable request.

**Supplementary Materials:**
Materials and Methods
Figures S1-S5
Table S1
Supplementary methods
References (##-##)

>>>>

---
output:
  word_document:
    fig_caption: yes
    reference_docx: ../text/word_template_science.docx
bibliography: ../text/global_proc.bib
csl: nature.csl
---

@@

**Abstract:**
Preserving coral reef functioning is a critical challenge of the 21st century. However, a lack of quantitative assessments of multiple functions across large spatial scales has hindered local and regional conservation efforts. We integrate empirically-parametrized bioenergetic models and global community surveys to quantify five key functions mediated by coral reef fishes. We show that functions exhibit critical trade-offs driven by varying community structures, such that no community can maximize all functions. Further, functions are locally dominated by few species, but worldwide, the identity of dominant species greatly varies; more than half of the 1,110 species in our dataset are functionally dominant on a local scale. Our results underline the need for a nuanced approach to coral reef conservation that considers multiple functions beyond the effect of standing stock biomass.

**Main Text:**
The flow of elements through biological communities fuels all ecosystems on earth [@Welti2017]. Harmful human dominance disrupts the natural functioning in contemperory ecosystems [@Ceballos2015].
Coral reefs represent a prime example of an ecosystem that is severely impacted by anthropogenic threats such as warming, and intensive fishing pressure [@Hughes2017; @Pauly2002].
Drastic declines in habitat quality and fish biomass have evoked serious concerns about the persistence of coral reef functioning [@Bellwood2019; @Williams2019; @Brandl2019front).
However, our knowledge on process-based functions - fluxes of elements and energy - is still in its infancy and evaluations of coral reef functioning typically rely on static proxies, such as live coral cover, standing stock biomass of reef fishes, or measures of diversity [e.g. @Cinner2020; @Mouillot2014; @Mora2011].
This sparsity of quantification of functions constitutes a severe limitation to efficient management of coral reef ecosystem functioning [@Brandl2019front].

Coral reef fishes play an essential role in coral reef functioning, because they contribute to a high proportion of consumer biomass and therefore represent essential vectors of fluxes of carbon (C), nitrogen (N) and phosphorus (P) through storage and recycling [@McIntyre2008; @Allgeier2014; @Barneche2014].
Ecosystem functions mediated by coral reef fishes include nutrient cycling, biomass production, herbivory , and piscivory (secondary consumption) (@Brandl2019front).
While the high diversity of coral reef fishes has inspired many studies with a focus on ecosystem functioning, only a handful of studies have attempted to quantify functions as continuous fluxes [@Brandl2019front]. Further, studies that do quantify functions mostly focused on biomass production (e.g., @Morais2019; @Morais2020glob) or fish excretion (e.g., @Allgeier2014), and suggest that functions rely on community structure. An assessment of multiple functions on a large scale is currently lacking.
Consequently, trade-offs amongs functions, their drivers, and their vulnerability to global stressors remain poorly understood in coral reef ecosystems [@Brandl2019front].

Here, we integrate biogeochemistry and community ecology to advance our understanding of the elemental fluxes that underpin reef fish functioning.
Using empirically-collected species-specific data on basic organismal processes and Bayesian phylogenetic models, we parameterize individual-level bioenergetic models to estimate five key ecosystem functions: N excretion, P excretion, biomass production, herbivory, and piscivory.
We apply these bioenergetic models to 9,118 reef fish transects across 585 sites worldwide (Table S1) to:
(1) quantify community-level reef fish functions and their trade-offs,
(2) extract the community- and species-level effects on these functions, and
(3) gauge the vulnerability of reef fish functioning in the Anthropocene.

@@

Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we demonstrate a predictably strong relationship between fish biomass and all five functions by running multivariate mixed effect Bayesian model (Fig. 1). Specifically, the effects of the log-transformed biomass on log-transformed functions were 0.932 (95\%CI: 0.929; 0.934), 1.051 (1.047; 1.056), 0.771 (0.764; 0.780), 0.940 (0.923; 0.957), and 0.668 (0.635; 0.702) for N excretion, P excretion, production, herbivory, and piscivory, respectively. We also incorporated sea surface temperature (sst) because it affects the metabolic and growth rates of fishes, which scales up to the community [@Barneche2019]. We found positive effects of sst on N excretion, production, and herbivory, and negligible effects of sst on P excretion and piscivory (table Sx).

The above-mentioned multivariate model also allowed us to estimate the correlations between functions, independent of the effects of biomass and sst.
Specifically, we estimated correlations between functions on three levels: the locality effects, the site effects, and the residual variations (Fig. 2, fig. Sx.
The correlations mostly overlapped across the three levels for a given pair of functions. We found consistently negative trade-offs between P excretion and N excretion, P excretion and biomass production. Further, there were slightly weaker negative correlations between piscivory and N excretion and between piscivory and herbivory. As a consequence of the negative correlations, a fish community cannot have high values for all functions simultaneously.

@@

We then questioned if functions tend to be driven by specific species across sites. We quantified the degree of dominance inside each community on the site-level and found that functions are on average charachterized by a high level of dominance, i.e. in most communities, a small proportion of species perform most of a function (Fig. 4a). As a comparison, we simulated random communities ranging from 5 to 50 species which yielded an average XX. The predicted average of the degree of dominance was x, x, x, x, and x for each function respectively.
We also calculated the proportion of species that is dominant in at least one transect (i.e. species that contribute more than a community in which each species contributes equally). We found that most species play a dominant role, and 70\% of all species contributed disproportionally to a function in at least one community (Fig. 4b).
Finally, we investigated how often species are dominant across their range and found that very few species are dominant throughout their range (Fig. 4c). Together, this means that functions inside communities tend to be driven by few dominant species, but that the identity of these dominant species varies highly across communities.

@@

By quantifying five key functions mediated by coral reef fishes, we demonstrate that ecosystem-wide functioning is mainly driven by fish biomass, yet biomass alone does not fully capture functional variability across communities.
While biomass is one of the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], our analyses suggests that biomass alone does not sufficiently characterize functioning.
We demonstrate strong trade-offs among the five functions, independent of biomass. Using biomass as a proxy for functioning can therefore mask fundamental differences in critical community-level functions.
Furthermore, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize all functions simultaneously.

Fish community structures drive the observed trade-offs among functions which can be explained by organismal physiology and life-history traits of the individuals inside communities [@Schiettekatte2020; @Schramski2015].
For example, we observed a clear trade-off between P excretion and biomass production which is largely driven by the age and trophic community structure.
Communities dominated by fishes with high trophic levels are charachterized high P excretion rates because predatory fishes have a P-rich diet (@Allgeier2014).
In contrast, biomass production is high in communities dominated by fishes with low trophic levels because herbivores tend to have higher growth rates [@Morais2018].
Moreover, communities dominated by mature fishes have high values of P excretion because growing fishes are more likely to retain P to form P-rich bones [@Schiettekatte2020; @Hood], while communities dominated by immature fishes maximize biomass production.
Furthermore, metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportionate contributions to functions that rely on rapid energetic turnover such as herbivory, piscivory, production, and N excretion [@Barneche2018; @Morais2020glob; @Brandl2019science].

Our results also reveal that functions consistently hinge on a few dominant species, but the identities of local, dominant species vary strongly across sites [@Lefcheck2019].
Locally, specific high-performing taxa may disproportionately impact rates of functioning at the community level due to high biomass or abundance [@Topor2019].
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
While few high-performing taxa indeed dominate functioning in each site, there are no species that are dominant across their entire range.
In addition, more than half of all species contributed disproportionally to a specific function in at least one site inside their range.
There are thus no widespread key species to target for conservation, rather, the functional role and conservation priority of species should be investigated at a local scale.

@@

Conserving biomass, diversity, and ecosystem functions are important objectives of most conservation initiatives [@Cinner2020].
While safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory.
However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards herbivore domination may result in the decline of P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions, yet our comprehension of
process-based functioning of coral reefs or what a “functional” coral reef really means is still in its infancy [@Brandl2019front].
To appropriately manage coral reef functioning with respect to local needs and stressors, future studies should seek to establish functional baselines for global coral reefs.
Until then, our study suggests that coral reef fish functions can be managed by enhancing standing stock biomass, protecting local key species, and promoting a balanced community structure.

@@

**References**
<div id="refs"></div>

**Acknowledgements:**
We thank the staff at CRIOBE, Moorea for field support. We would also like to thank Jérémy Carlot, Beverly French, Titouan Roncin, Yann Lacube, Camille Gache, Gabrielle Martineau, Kailey Bissell, Benoit Espiau, Calvin Quigley, Kaitlyn Landfield and Tommy Norin for their help in the field, Sophie Schiettekatte for proof-reading the manuscript, and Guillemette de Sinéty and Jérémy Wicquart for their contribution to otolith analyis.
This research was funded by the BNP Paribas Foundation (Reef Services Project) and the French National Agency for Scientific Research (ANR; REEFLUX Project; ANR‐17‐CE32‐0006). This research is product of the SCORE-REEF group funded by the Centre de Synthèse et d’Analyse sur la Biodiversité (CESAB) of the Foundation pour la Recherche sur la Biodiversité (FRB) and the Agence Nationale de la Biodiversité (AFB). VP was supported by the Institut Universitaire de France (IUF) and JMC was supported by a Make Our Planet Great Again Postdoctoral Grant (mopga‐pdf‐0000000144).
**Author contributions:**
NMDS and VP conceived the idea and NMDS, VP, SJB, and JMC designed methodology; NMDS, JMC, SJB, AM, FM, VP, KSM, JEA and DEB collected the data; All authors shared existing data. NMDS analyzed the data and led the writing of the manuscript. All authors contributed significantly to the drafts and approved the final version for publication.
**Competing interests:** None declared.
**Data and materials availability:**
All data and code to reproduce figures are available online through GitHub (https://github.com/nschiett/global_proc) and figshare (https://figshare.com/s/f789aec2c20492c4f0f9). All data on individual fish traits are available from the corresponding author on reasonable request.

**Figure legends**

@@

### 2021-09-09 text/methods.Rmd
methods edits

## 2. Quantification of functions
\noindent For each transect, we estimated five key process-based functions mediated by fishes: nitrogen excretion rate (gN/day/m^2^), phosphorus excretion rate (gP/day/m^2^), production of biomass through growth (gC/day/m^2^), herbivory, i.e. ingestion rate of macrophytes (gC/day/m^2^), and piscivory, i.e. ingestion rate of fishes (g/day/m^2^) [@Brandl2019front]. These five functions were estimated for each transect using individual-based bioenergetic models predicting fluxes of carbon (C), nitrogen (N), and phosphorus (P) (e.g. daily C intake rates, N and P excretion rates, growth rates) [@Schiettekatte2020].
This bioenergetic model framework integrates elements of metabolic theory, stoichiometry, and flexible elemental limitation [@Schiettekatte2020]. We quantified the input parameters, including elements of metabolism, growth, and diet and body stoichiometry, for all 1110 species through the integration of empirical data, data synthesis, and Bayesian phylogenetic models (See supplementary methods). We then ran a unique bioenergetic model for each combination of species identity, body size, and sea surface temperature (n = 30668) to get the contribution of each individual to each function in each transect. Finally, we summarized functions on the community level by summing up all individual contributions inside a transect and deviding the sum by the surface area. Each function is thus expressed as dry mass (of C, N, or P) per day per square meter. We note that N excretion, P excretion, and biomass production include contributions of all fishes, whereas herbivory and piscivory are carried out by a subset of the community, with respect to their trophic guild as defined by @Parravicini2020. To reduce the occurrence of misclassification of herbivores and piscivores, we categorized a species as a herbivore or piscivore if it had both the highest probability to be classified in that trophic group and this probability was more than 0.5, based on the probability scores of trophic guilds for a global fish species database that defines trophic guilds based on empirical data using a quantitative, unbiased, and fully reproducible framework [@Parravicini2020]. Further, as a comparison, we quantified herbivory and piscivory rates using two alternative trophic guild classifications based on expert opinion [@Mouillot2014; @Parravicini2020] (Fig. S5). Both the herbivory and piscivory rates are congruent with the expert opinion trophic guild classifications.
Finally, we estimated multifunction, i.e. one measure that combines all five functions by taking the geometric average of the five standardized functions. We used the geometric mean because functions are dependent on each other and may vary considerably.

@@

\noindent Finally, to investigate the effect of the community structure while still accounting for the effects of standing biomass and sea surface temperature, we fitted a mixed effect multivariate model similar to the model specified above, but adding all community structure variables:

@@

All Bayesian models were fitted using the R package *brms* (ref) which uses Stan, a C++ package for performing full Bayesian inference. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by using four chains of 2,000 samples, including 1,000 samples as a warm‐up, meaning that a total of 4,000 samples were used to estimate posterior distributions. The convergence and fit of the models were verified by examining the Rhat, parameter trace plots, and posterior prediction plots. Posterior predictive checks for are provided in Supporting Information xxxx.

## 5. Species dominance and contributions to functions
We quantified the relative contribution of each species to each function for all transects as followed:
<!-- \begin{equation} -->
$$	\textrm{contribution}_{f,i,j} = \frac{F_{f,i,j}}{\sum F_{f,j}},$$
<!-- 	\label{con} -->
<!-- \end{equation} -->
where i is a certain species, j is a transect, F is the value of function f.

Then, we quantified the degree of species dominance per function for each transect.
We did this by first ranking species according to their contribution to function, followed by quantifying the cumulative contributions of species to functions. Then, we used the area under the species accumulation curve as a measure for the degree of dominance.
Specifically, the degree of dominance (DD) was calculated as followed:
<!-- \begin{equation} -->
$$	\textrm{DD} = \frac{A - A_{min}}{A_{max} - A_{min}},$$
<!-- 	\label{dd} -->
<!-- \end{equation} -->
where $A$ is the area under the curve, $A_{min}$ is the theoretical area under the curve where each species has an equal contribution to a certain function, $A_max$ is the theoretical area under the curve where one species performs the entire function. They are quantified as:
<!-- \begin{equation} -->
$$	\textrm{A}_{min} = \frac{R^2 - 1}{2 R},$$
<!-- 	\label{Amin} -->
<!-- \end{equation} -->

@@

Finally, we quantified the frequency of dominance per species, i.e. the number of transects in which a species is dominant for a given function divided by the total number of transects in which that species is observed. A species is considered dominant for a certain function in a given transect if their contribution is higher than 1/R, i.e. they contribute more than the situation in which each species contributes equally to a certain function.

## 6. Vulnerability to fishing and climate change
First, we estimated or extracted two measures of vulnerability per species: vulnerability to climate change and vulnerability to fishing pressure [@Graham2011]. For species' vulnerability to climate change, we solely focus on their vulnerability to the loss of live corals. Vulnerability to climate change induced coral loss is related to diet specialization, habitat specialization for live coral, and body size [@Graham2011]. Specifically, we used a score for climate change vulnerability for 134 species found in Graham et al. (2011) [@Graham2011] and fitted a Bayesian mixed effect predictive model that relates these vulnerability scores with the log-transformed maximum size of fish (extracted from Fishbase @Froese2018), the dependence on coral for food (3 categories: not dependent, facultative corallivore, and obligate corallivore), and dependence on coral for habitat (2 categories: dependent vs. not dependent) [@Coker2014; @Cole2008]. We also included a random effect for family:
$$ y_i \sim normal(\mu_i, \sigma)$$
$$
\mu_i = (\beta0 +  \delta_{family}) + \beta1 log(sizemax) + \beta2 HD + \beta3 FD  ,
$$
with $y_i$ the vulnerability score of species i, $\mu_i$ the average predicted vulnerability score, sizemax the maximum length of the species, HD de dependence for habitat, and FD the depedendence for food.
To verify the fit of the model we inspected the posterior predictive plot, which indicated a good fit. Further, the model had a Bayesian R2 of 0.97. We then used this model to extrapolate the vulnerability measure to all 1110 species in our dataset. We provide the model parameters in supplementary table xx. \par For species' vulnerability to fishing, we extracted the index from Cheung et al. (2005) [@Cheung2005].
Next, we calculated vulnerability scores per function on the community level by averaging the species-level scores weighted by the contributions to function of species.
Finally, we fitted a mixed model to predict the average vulnerability to climate change and fishing for each function.

>>>>

## 2. Quantification of functions
\noindent For each transect, we estimated five key process-based functions mediated by fishes: nitrogen excretion rate (gN/day/m^2^), phosphorus excretion rate (gP/day/m^2^), production of biomass through growth (gC/day/m^2^), herbivory, i.e. ingestion rate of macrophytes (gC/day/m^2^), and piscivory, i.e. ingestion rate of fishes (g/day/m^2^) [@Brandl2019front]. These five functions were estimated for each transect using individual-based bioenergetic models predicting fluxes of carbon (C), nitrogen (N), and phosphorus (P) (e.g. daily C intake rates, N and P excretion rates, growth rates) [@Schiettekatte2020].
This bioenergetic model framework integrates elements of metabolic theory, stoichiometry, and flexible elemental limitation [@Schiettekatte2020]. We quantified the input parameters, including elements of metabolism, growth, and diet and body stoichiometry, for all 1110 species through the integration of empirical data, data synthesis, and Bayesian phylogenetic models (See supplementary methods). We then ran a unique bioenergetic model for each combination of species identity, body size, and sea surface temperature (n = 30668) to get the contribution of each individual to each function in each transect. Finally, we summarized functions on the community level by summing up all individual contributions inside a transect and deviding the sum by the surface area. Each function is thus expressed as dry mass (of C, N, or P) per day per square meter. We note that N excretion, P excretion, and biomass production include contributions of all fishes, whereas herbivory and piscivory are carried out by a subset of the community, with respect to their trophic guild as defined by @Parravicini2020. To reduce the occurrence of misclassification of herbivores and piscivores, we categorized a species as a herbivore or piscivore if it had both the highest probability to be classified in that trophic group and this probability was more than 0.5, based on the probability scores of trophic guilds for a global fish species database that defines trophic guilds based on empirical data using a quantitative, unbiased, and fully reproducible framework [@Parravicini2020]. Further, as a comparison, we quantified herbivory and piscivory rates using two alternative trophic guild classifications based on expert opinion [@Mouillot2014; @Parravicini2020] (Fig. S5). Both the herbivory and piscivory rates are congruent with the expert opinion trophic guild classifications.
Finally, we estimated multifunction, i.e. one measure that combines all five functions by taking the geometric average of the five functions ( nomalized to range between zero and 100). We used the geometric mean because functions are dependent on each other and may vary considerably.

@@

\noindent Finally, to investigate the effect of the community structure while still accounting for the effects of standing biomass and sst, we fitted a mixed effect multivariate model similar to the model specified above, but adding all community structure variables:

@@

All Bayesian models were fitted using the R package *brms* [@Burkner2017] which uses Stan, a C++ package for performing full Bayesian inference [@Carpenter2017]. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by using four chains of 2,000 samples, including 1,000 samples as a warm‐up, meaning that a total of 4,000 samples were used to estimate posterior distributions. The convergence and fit of the models were verified by examining the Rhat, parameter trace plots, and posterior prediction plots (Fig Sx).

## 5. Species dominance and contributions to functions
We quantified the relative contribution of each species to each function for all sites as followed:
<!-- \begin{equation} -->
$$	\textrm{contribution}_{f,i,j} = \frac{F_{f,i,j}}{\sum F_{f,j}},$$
<!-- 	\label{con} -->
<!-- \end{equation} -->
where i is a certain species, j is a site, F is the value of function f.

Then, we quantified the degree of species dominance per function for each site.
We did this by first ranking species according to their contribution to function, followed by quantifying the cumulative contributions of species to functions. Then, we used the area under the species accumulation curve as a measure for the degree of dominance.
Specifically, the degree of dominance (DD) was calculated as followed:
<!-- \begin{equation} -->
$$	\textrm{DD} = \frac{A - A_{min}}{A_{max} - A_{min}},$$
<!-- 	\label{dd} -->
<!-- \end{equation} -->
where $A$ is the area under the curve, $A_{min}$ is the theoretical area under the curve where each species has an equal contribution to a certain function, $A_max$ is the theoretical area under the curve where one species performs the entire function. They are quantified as:
<!-- \begin{equation} -->
$$	\textrm{A}_{min} = \frac{R^2 - 1}{2 R},$$
<!-- 	\label{Amin} -->
<!-- \end{equation} -->

@@

Finally, we quantified the frequency of dominance per species, i.e. the number of sites in which a species is dominant for a given function divided by the total number of sites in which that species is observed. A species is considered dominant for a certain function in a given site if their contribution is higher than 1/R, i.e. they contribute more than the situation in which each species contributes equally to a certain function.
